The \dealii Library, Version 9.6

### 2024-09-25 9.6/paper.tex

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item \dealii{} now requires and makes use of the C++17 language standard.
  \item We have continued to make progress in supporting simplex and
    mixed meshes -- mesh types that \dealii{} has traditionally not
    supported at all. Specifically, the current release uses
    better strategies for refinement of tetrahedra that result in
    better-shaped child cells. It also contains
    support for cubic finite elements on simplices.
   \item The \texttt{FE\_NedelecSZ} class that contains our 
    implementation of the N\'ed\'elec element using the orientation 
    scheme of \cite{Zag06} now supports the computation of hanging 
    node constraints for locally refined, hexahedral meshes. 
    That is, the implementation of hanging node constraints now
    correctly addresses the sign-conflict that arises for $hp$ quadrilateral and 
    hexahedral N\'ed\'elec elements; 
    see \cite{Kinnewig2024} for details. 
    The other implementation of N\'ed\'elec elements, in the \texttt{FE\_NedelecSZ} class, already
    implements hanging node constraints; therefore, there is no longer
    a difference for the user between the two classes as far as
    constraints are concerned. 
    However, the special case where, in 3D, more than four cells with 
    different refinement levels share a common edge is not covered yet.
  \item The \texttt{AffineConstraints} class stores and processes
    constraints on degrees of freedom in \dealii{}. Such constraints
    can be of the (homogeneous) form $x_3 = \frac 12 x_{14} + \frac 12 x_{15}$ as is
    common when using hanging node constraints (here, $x_3$ could
    represent the value of the solution at a hanging node, and $x_{14}$
    and $x_{15}$ are the values of the two adjacent degrees of freedom on
    the parent edge that contains the hanging node); or they can be of
    the (inhomogeneous) form $x_{12}=42$ as is common when using Dirichlet boundary
    conditions.

@@

In the current release, we have put substantial work into this switch,
as outlined below. As this is not yet finished, it is important to note that 
all these features are optional and their individual availability
depends on the installed \trilinos{} packages. 
As an example, a \dealii{} build will pass with a \trilinos{} installation 
without \texttt{Ifpack2}, but its preconditioners will not be included.
In the following list, we omit the common namespace prefix
\texttt{LinearAlgebra::} on all mentioned
\texttt{TpetraWrappers} symbols for readability:
\begin{itemize}
  \item The \texttt{TpetraWrappers::Vector} class has been overhauled to also allow specifying
    a memory space, and the \texttt{TpetraWrappers::BlockVector} class has been added.
  \item \texttt{TpetraWrappers::SparseMatrix} and \texttt{TpetraWrappers::SparsityPattern} have been implemented mirroring
    the functionality of the \texttt{TrilinosWrappers::SparseMatrix} \\ and \texttt{TrilinosWrappers::SparsityPattern} classes.
  \item Various incomplete factorization and relaxation preconditioners (from \texttt{Ifpack2}) have been wrapped,
    mirroring the existing classes in \texttt{LinearAlgbera::TrilinosWrappers} as closely as possible.
  \item Two new preconditioner variants \texttt{TpetraWrappers::PreconditionL1Jacobi} and \\
    \texttt{TpetraWrappers::PreconditionL1GaussSeidel}, based on~\cite{BFKY2011}, have been added.
  \item \texttt{TpetraWrappers::SolverDirectKLU2} is now available as a direct solver.
\end{itemize}
Some functionality in these wrapper classes is still missing, 
most noticeably wrappers for the algebraic-multigrid preconditioner \texttt{MueLu}
and the iterative solvers from \texttt{Belos}.
However, the wrapped \texttt{Ifpack2} preconditioners can already be used with the iterative solvers of \dealii{}.

@@

The design goal was to introduce as few changes as possible for the user to allow for an easy transition to the
\texttt{TpetraWrappers} classes.
However, based on the difference between \epetra{} and \tpetra{}, the following changes were necessary:
\begin{itemize}
  \item The constructor for the class \texttt{TrilinosWrappers::MPI::Vector} only required the \texttt{IndexSet locally\_owned},
    which describes the set of indices locally owned by the current rank.
    However, if one wants to create a vector that has read or write access to non-locally owned indices,  the constructor of the
    \texttt{TpetraWrappers::Vector} requires the \texttt{IndexSet locally\_owned}, the \texttt{IndexSet locally\_relevant}, and
    a boolean flag, whether to initialize the vector in the read-only or the write-only state.
    This interface is in fact close to the one of \dealii{}'s own parallel vector class.
  \item A vector created without providing the \texttt{IndexSet locally\_relevant} is purely local and cannot access non-local indices.
    Such a vector cannot be copied to a vector that can access non-local indices, as the \texttt{IndexSet locally\_relevant} must
    be provided at creating the vector object. 
  \item Some of parameters of existing solver and preconditioner are
    not available in the \tpetra{} wrappers,
    such that the corresponding \texttt{AdditionalData} objects are not identical. 
    Since there are many preconditioners we will not list each individual change but instead refer 
    to our Doxygen documentation. 
  \item The incomplete Cholesky (IC) factorization preconditioner is not available 
   in \tpetra{}.
\end{itemize}

@@

% This is up-to-date as of 2024-07-23 11:52PM CEST
\begin{quote}
Laryssa      Abdala,
Mathias      Anselmann,
Abbas        Ballout,
Maximilian   Bergbauer,
Julian       Brotz,
Marco        Feder,
Niklas       Fehn,
Menno        Fraters,
Quang        Hoang,
Vladimir     Ivannikov,
Tao          Jin,
Yimin        Jin,
Sebastian    Kinnewig,
Paras        Kumar,
Sébastien    Loriot,
Nils         Much,
Abdullah     Mujahid,
Bob          Myhill,
Paul A.      Patience,
Luz          Paz,
Laura        Prieto Saavedra,
Sebastian    Proell,
Hendrik      Ranocha,
Johannes     Resch,
Andreas      Ritthaler,
Malik        Scheifinger,
David        Schneider,
%Magdalena    Schreter-Fleischhacker,
Richard      Schussnig,
Nils         Schween,
Kyle         Schwiebert,
Simranjeet   Singh,
Simon        Sticko,
Dominik      Still,
%Jan Philipp  Thiele,
Thierry      Thomas,
Vinayak      Vijay,
Ivy          Weber,
Simon        Wiesheier,
Chengjiang   Yin,
%Vladimir     Yushutin,
Stefano      Zampini.
\end{quote}
Their contributions are much appreciated!

@@

This work used [resource-name] at [resource provider] through allocation [allocation number] from the Advanced Cyberinfrastructure Coordination Ecosystem: Services \& Support (ACCESS) program, which is supported by National Science Foundation grants \#2138259, \#2138286, \#2138307, \#2137603, and \#2138296. See \cite{Boerner2023}.

>>>>

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item \dealii{} now requires and makes use of the C++17 language standard.
  \item We have continued to make progress in supporting simplex and
    mixed meshes -- mesh types that \dealii{} has traditionally not
    supported at all. Specifically, the current release uses
    better strategies for refinement of tetrahedra that result in
    better-shaped child cells. It also contains
    support for cubic finite elements on simplices.
   \item The \texttt{FE\_NedelecSZ} class that contains our 
    implementation of the N\'ed\'elec element using the orientation 
    scheme of \cite{Zag06} now supports the computation of hanging 
    node constraints for locally refined, hexahedral meshes. 
    That is, the implementation of hanging node constraints now
    correctly addresses the sign-conflict that arises for $hp$ quadrilateral and 
    hexahedral N\'ed\'elec elements; 
    see \cite{Kinnewig2024} for details. 
    The other implementation of N\'ed\'elec elements, in the \texttt{FE\_NedelecSZ} class, already
    \todo[inline]{JPT: I think this should be \texttt{FE\_Nedelec}, without SZ?}
    implements hanging node constraints; therefore, there is no longer
    a difference for the user between the two classes as far as
    constraints are concerned. 
    However, the special case where, in 3D, more than four cells with 
    different refinement levels share a common edge is not covered yet.
  \item The \texttt{AffineConstraints} class stores and processes
    constraints on degrees of freedom in \dealii{}. Such constraints
    can be of the (homogeneous) form $x_3 = \frac 12 x_{14} + \frac 12 x_{15}$ as is
    common when using hanging node constraints (here, $x_3$ could
    represent the value of the solution at a hanging node, and $x_{14}$
    and $x_{15}$ are the values of the two adjacent degrees of freedom on
    the parent edge that contains the hanging node); or they can be of
    the (inhomogeneous) form $x_{12}=42$ as is common when using Dirichlet boundary
    conditions.

@@

In the current release, we have put substantial work into this switch,
as outlined below. As this is not yet finished, it is important to note that 
all these features are optional and their individual availability
depends on the installed \trilinos{} packages. 
As an example, a \dealii{} build will pass with a \trilinos{} installation 
without \texttt{Ifpack2}, but its preconditioners will not be included.
In the following list, we omit the common namespace prefix
\texttt{LinearAlgebra::} on all mentioned
\texttt{TpetraWrappers} symbols for readability:
\begin{itemize}
  \item The \texttt{TpetraWrappers::Vector} class has been overhauled to also allow specifying
    a memory space, and the \texttt{TpetraWrappers::BlockVector} class has been added.
  \item \texttt{TpetraWrappers::SparseMatrix} and \texttt{TpetraWrappers::SparsityPattern} have been implemented mirroring
    the functionality of the \texttt{TrilinosWrappers::SparseMatrix} \\ and \texttt{TrilinosWrappers::SparsityPattern} classes.
  \item Various incomplete factorization and relaxation preconditioners (from \texttt{Ifpack2}) have been wrapped,
    mirroring the existing classes in \texttt{LinearAlgbera::TrilinosWrappers} as closely as possible.
  \item Two new preconditioner variants \texttt{TpetraWrappers::PreconditionL1Jacobi} and \\
    \texttt{TpetraWrappers::PreconditionL1GaussSeidel}, based on~\cite{BFKY2011}, have been added.
  \item \texttt{TpetraWrappers::SolverDirectKLU2} is now available as a direct solver.
\end{itemize}
Some functionality in these wrapper classes is still missing, 
most noticeably wrappers for the algebraic-multigrid preconditioner \texttt{MueLu}
and the iterative solvers from \texttt{Belos}.
However, the wrapped \texttt{Ifpack2} preconditioners can already be used with the iterative solvers of \dealii{}.
To do this you will need to explicitly specify the vector type parameter,
e.g.\\ \texttt{SolverCG<LinearAlgebra::TpetraWrappers::Vector<double, MemorySpace::Default> >}

@@

The design goal was to introduce as few changes as possible for the user to allow for an easy transition to the
\texttt{TpetraWrappers} classes.
However, based on the difference between \epetra{} and \tpetra{}, the following changes were necessary:
\begin{itemize}
  \item The constructor for the class \texttt{TrilinosWrappers::MPI::Vector} only required the \texttt{IndexSet locally\_owned},
    which describes the set of indices locally owned by the current rank.
    However, if one wants to create a vector that has read or write access to non-locally owned indices,  the constructor of the
    \texttt{TpetraWrappers::Vector} requires the \texttt{IndexSet locally\_owned}, the \texttt{IndexSet locally\_relevant}, and
    a boolean flag, whether to initialize the vector in the read-only or the write-only state.
    This interface is in fact close to the one of \dealii{}'s own parallel vector class.
  \item A vector created without providing the \texttt{IndexSet locally\_relevant} is purely local and cannot access non-local indices.
    Such a vector cannot be copied to a vector that can access non-local indices, as the \texttt{IndexSet locally\_relevant} must
    be provided at creating the vector object. 
  \item Some parameters of existing solver and preconditioner are
    not available in the \tpetra{} wrappers,
    such that the corresponding \texttt{AdditionalData} objects are not identical. 
    Since there are many preconditioners we will not list each individual change but instead refer 
    to our Doxygen documentation. 
  \item The incomplete Cholesky (IC) factorization preconditioner is not available 
   in \tpetra{}.
\end{itemize}

@@

% This is up-to-date as of 2024-07-23 11:52PM CEST
\begin{quote}
Laryssa      Abdala,
Mathias      Anselmann,
Abbas        Ballout,
Maximilian   Bergbauer,
Julian       Brotz,
Marco        Feder,
Niklas       Fehn,
Menno        Fraters,
Quang        Hoang,
Vladimir     Ivannikov,
Tao          Jin,
Yimin        Jin,
%Sebastian    Kinnewig,
Paras        Kumar,
Sébastien    Loriot,
Nils         Much,
Abdullah     Mujahid,
Bob          Myhill,
Paul A.      Patience,
Luz          Paz,
Laura        Prieto Saavedra,
Sebastian    Proell,
Hendrik      Ranocha,
Johannes     Resch,
Andreas      Ritthaler,
Malik        Scheifinger,
David        Schneider,
%Magdalena    Schreter-Fleischhacker,
Richard      Schussnig,
Nils         Schween,
Kyle         Schwiebert,
Simranjeet   Singh,
Simon        Sticko,
Dominik      Still,
%Jan Philipp  Thiele,
Thierry      Thomas,
Vinayak      Vijay,
Ivy          Weber,
Simon        Wiesheier,
Chengjiang   Yin,
%Vladimir     Yushutin,
Stefano      Zampini.
\end{quote}
Their contributions are much appreciated!

@@

This work used [resource-name] at [resource provider] through allocation [allocation number] from the Advanced Cyberinfrastructure Coordination Ecosystem: Services \& Support (ACCESS) program, which is supported by National Science Foundation grants \#2138259, \#2138286, \#2138307, \#2137603, and \#2138296. See \cite{Boerner2023}.
\todo[inline]{Missing ACCESS information}
